Racial and Ethnic Differences in ETI: Outcomes

### 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
much faster outcomes processing (what was I doing?!)

>>>>

<!-- ### Full cohort -->

<!-- ```{r} -->
<!-- # Fit mixed model with change point -->
<!-- change_mod <- glmer( -->
<!--   PEx ~ Years * Race + slope_change_years * Race + age_group + -->
<!--     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa + -->
<!--     (1 | eDWID), -->
<!--   family = "binomial", -->
<!--   data = encounter -->
<!-- ) -->
<!-- # Model plot -->
<!-- pop_dat <- data.frame(expand.grid( -->
<!--   Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)", -->
<!--   Insurance = "Private", prior_modulator = "Yes", -->
<!--   isOnEnzymes = "Yes", pseudomonasaeruginosa = "No" -->
<!-- )) -->
<!-- pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years -->
<!-- pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T) -->
<!-- pop_dat$pred <- pred$fit -->
<!-- pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit) -->
<!-- pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit) -->
<!-- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) + -->
<!--   geom_line() + -->
<!--   xlab("Years from ETI Start") + -->
<!--   ylab("gli_fev1_ppred_rn") + -->
<!--   theme_classic() -->
<!-- # Model summary -->
<!-- change_mod %>% -->
<!--   gtsummary::tbl_regression(intercept = T) %>% -->
<!--   as_gt() %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- # Table of predicted values -->
<!-- pop_dat %>% -->
<!--   select(-slope_change_years) %>% -->
<!--   rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>% -->
<!--   gt(caption = "Predicted Values at Various Timepoints") %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- # Contrasts -->
<!-- aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) -->
<!-- aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0) -->
<!-- hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0) -->
<!-- hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) -->
<!-- nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
<!-- nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
<!-- other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0) -->
<!-- other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1) -->
<!-- contr <- rbind( -->
<!--   "Black/AA Pre-ETI" = aa_pre, -->
<!--   "Black/AA Post-ETI" = aa_pre + aa_change, -->
<!--   "H/L Pre-ETI" = hl_pre, -->
<!--   "H/L Post-ETI" = hl_pre + hl_change, -->
<!--   "NHW Pre-ETI" = nhw_pre, -->
<!--   "NHW Post-ETI" = nhw_pre + nhw_change, -->
<!--   "Other Pre-ETI" = other_pre, -->
<!--   "Other Post-ETI" = other_pre + other_change -->
<!-- ) -->
<!-- change_mod_slopes <- glht(change_mod, linfct = contr) -->
<!-- change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha()) -->
<!-- change_mod_slopes_ci$confint %>% -->
<!--   as.data.frame() %>% -->
<!--   gt(caption = "Slopes by Race", rownames_to_stub = T) %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- ``` -->

<!-- ### Age < 18 at ETI initiation -->

<!-- The following models were not adjusted for age group. -->

<!-- ```{r} -->
<!-- # Fit mixed model with change point -->
<!-- change_mod <- lmer( -->
<!--   gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance + -->
<!--     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID), -->
<!--   data = encounter[encounter$age_eti_group == "[-Inf,18)", ] -->
<!-- ) -->
<!-- # Model plot -->
<!-- pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T) -->
<!-- pop_dat$pred <- pred$fit -->
<!-- pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit) -->
<!-- pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit) -->
<!-- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) + -->
<!--   geom_line() + -->
<!--   xlab("Years from ETI Start") + -->
<!--   ylab("gli_fev1_ppred_rn") + -->
<!--   theme_classic() -->
<!-- # Model summary -->
<!-- change_mod %>% -->
<!--   gtsummary::tbl_regression(intercept = T) %>% -->
<!--   as_gt() %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- # Table of predicted values -->
<!-- pop_dat %>% -->
<!--   select(-slope_change_years) %>% -->
<!--   rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>% -->
<!--   gt() %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- # Contrasts -->
<!-- aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) -->
<!-- aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0) -->
<!-- hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0) -->
<!-- hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) -->
<!-- nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
<!-- nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
<!-- other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0) -->
<!-- other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1) -->
<!-- contr <- rbind( -->
<!--   "Black/AA Pre-ETI" = aa_pre, -->
<!--   "Black/AA Post-ETI" = aa_pre + aa_change, -->
<!--   "H/L Pre-ETI" = hl_pre, -->
<!--   "H/L Post-ETI" = hl_pre + hl_change, -->
<!--   "NHW Pre-ETI" = nhw_pre, -->
<!--   "NHW Post-ETI" = nhw_pre + nhw_change, -->
<!--   "Other Pre-ETI" = other_pre, -->
<!--   "Other Post-ETI" = other_pre + other_change -->
<!-- ) -->
<!-- change_mod_slopes <- glht(change_mod, linfct = contr) -->
<!-- change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha()) -->
<!-- change_mod_slopes_ci$confint %>% -->
<!--   as.data.frame() %>% -->
<!--   gt(caption = "Slopes by Race", rownames_to_stub = T) %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- ``` -->

<!-- ### Age >= 18 at ETI initiation -->

<!-- The following models were not adjusted for age group. -->

<!-- ```{r} -->
<!-- # Fit mixed model with change point -->
<!-- change_mod <- lmer( -->
<!--   gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance + -->
<!--     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID), -->
<!--   data = encounter[encounter$age_eti_group == "[18, Inf)", ] -->
<!-- ) -->
<!-- # Model plot -->
<!-- pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T) -->
<!-- pop_dat$pred <- pred$fit -->
<!-- pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit) -->
<!-- pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit) -->
<!-- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) + -->
<!--   geom_line() + -->
<!--   xlab("Years from ETI Start") + -->
<!--   ylab("gli_fev1_ppred_rn") + -->
<!--   theme_classic() -->
<!-- # Model summary -->
<!-- change_mod %>% -->
<!--   gtsummary::tbl_regression(intercept = T) %>% -->
<!--   as_gt() %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- # Table of predicted values -->
<!-- pop_dat %>% -->
<!--   select(-slope_change_years) %>% -->
<!--   rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>% -->
<!--   gt() %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- # Contrasts -->
<!-- aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) -->
<!-- aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0) -->
<!-- hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0) -->
<!-- hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) -->
<!-- nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
<!-- nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
<!-- other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0) -->
<!-- other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1) -->
<!-- contr <- rbind( -->
<!--   "Black/AA Pre-ETI" = aa_pre, -->
<!--   "Black/AA Post-ETI" = aa_pre + aa_change, -->
<!--   "H/L Pre-ETI" = hl_pre, -->
<!--   "H/L Post-ETI" = hl_pre + hl_change, -->
<!--   "NHW Pre-ETI" = nhw_pre, -->
<!--   "NHW Post-ETI" = nhw_pre + nhw_change, -->
<!--   "Other Pre-ETI" = other_pre, -->
<!--   "Other Post-ETI" = other_pre + other_change -->
<!-- ) -->
<!-- change_mod_slopes <- glht(change_mod, linfct = contr) -->
<!-- change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha()) -->
<!-- change_mod_slopes_ci$confint %>% -->
<!--   as.data.frame() %>% -->
<!--   gt(caption = "Slopes by Race", rownames_to_stub = T) %>% -->
<!--   fmt_number(decimals = 3) -->
<!-- ``` -->
